Persistence of seed mixture species: <br> Analysis of Bauer et al. (unpublished) Field experiment

### 2022-11-14 R/model_persistence_markdown.Rmd
Update model_persistence_markdown.Rmd

```{r}
sites <- read_csv(here("data", "processed", "data_processed_sites.csv"),
                  col_names = TRUE, na = c("na", "NA", ""), col_types =
                    cols(
                      .default = "?",
                      plot = "f",
                      site = "f",
                      sand_ratio = "f",
                      substrate_depth = "f",
                      target_type = col_factor(levels = c(
                        "dry_grassland", "hay_meadow"
                      )),
                      seed_density = "f",
                      exposition = col_factor(levels = c(
                        "north", "south"
                      )),
                      survey_year = "d"
                    )) %>%
  ### Exclude data of seed mixtures
  filter(survey_year != "seeded") %>%
  pivot_longer(cols = c(B, C), names_to = "index", values_to = "n",
               names_transform = as.factor) %>%
  filter(index = C) %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, sep = " "),
    botanist_year = factor(botanist_year),
    id = factor(id)
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
  )
```

@@

```{r echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Sand ratio [vol%] (Data of 2021)")
plot2 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Substrate depth [cm] (Data of 2021)")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Target type (Data of 2021)")
plot4 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Seed density [g/m²] (Data of 2021)")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Exposition (Data of 2021)")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Blocks (Data of 2021)")
plot4 <- ggplot(sites, aes(y = n, x = botanist_year)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Botanists and survey year") +
  theme(axis.text.x = element_text(angle = 90))
(plot1 + plot2) / (plot3 + plot4)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

@@

```{r}
iter = 10000
chains = 4
thin = 2
priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0.1, 1)", class = "b", coef = "sand_ratio25"),
  set_prior("normal(0.2, 1)", class = "b", coef = "sand_ratio50"),
  set_prior("normal(0.1, 1)", class = "b", coef = "expositionsouth"),
  set_prior("normal(0.1, 1)", class = "b", coef = "survey_year_fct2019"),
  set_prior("normal(0.2, 1)", class = "b", coef = "survey_year_fct2020"),
  set_prior("normal(0.3, 1)", class = "b",coef = "survey_year_fct2021"),
  set_prior("cauchy(0, 1)", class = "sigma")
)
```

@@

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sd_botanist_year__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sd_botanist_year__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- nuts_params(m_1)
hmc_diagnostics2 <- nuts_params(m_2)
y <- sites$n
yrep1 <- posterior_predict(m_1, draws = 500)
yrep2 <- posterior_predict(m_2, draws = 500)
loo1 <- loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

```{r}
prior_summary(m_2, all = FALSE)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

@@

```{r}
draws2
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
)
mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
)
```

>>>>

```{r}
sites <- read_csv(
  here("data", "processed", "data_processed_sites_temporal.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
      .default = "?",
      plot = "f",
      site = "f",
      sand_ratio = "f",
      substrate_depth = "f",
      target_type = col_factor(levels = c(
        "dry_grassland", "hay_meadow"
      )),
      seed_density = "f",
      exposition = col_factor(levels = c(
        "north", "south"
      )),
      survey_year = "c"
    )
  ) %>%
  ### Exclude data of seed mixtures
  filter(survey_year != "seeded") %>%
  pivot_longer(cols = c(B, C), names_to = "index", values_to = "n",
               names_transform = as.factor) %>%
  filter(index == "B" & presabu == "presence") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    id = factor(id)
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, n
  )
```

@@

```{r echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Sand ratio [vol%] (Data of 2021)")
plot2 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Substrate depth [cm] (Data of 2021)")
plot3 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Target type (Data of 2021)")
plot4 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Seed density [g/m²] (Data of 2021)")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Exposition (Data of 2021)")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Blocks (Data of 2021)")
(plot1 + plot2) / (plot3)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

@@

```{r}
iter = 10000
chains = 4
thin = 2
priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(-0.025, 1)", class = "b", coef = "sand_ratio25"),
  set_prior("normal(-0.050, 1)", class = "b", coef = "sand_ratio50"),
  set_prior("normal(0.050, 1)", class = "b", coef = "expositionsouth"),
  set_prior("normal(-0.025, 1)", class = "b", coef = "survey_year_fct2019"),
  set_prior("normal(-0.050, 1)", class = "b", coef = "survey_year_fct2020"),
  set_prior("normal(-0.075, 1)", class = "b", coef = "survey_year_fct2021"),
  set_prior("cauchy(0, 1)", class = "sigma")
)
```

@@

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- nuts_params(m_1)
hmc_diagnostics2 <- nuts_params(m_2)
y <- sites$n
yrep1 <- posterior_predict(m_1, draws = 500)
yrep2 <- posterior_predict(m_2, draws = 500)
loo1 <- loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

```{r}
prior_summary(m_1, all = FALSE)
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

@@

```{r}
draws1
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
)
mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
)
```
